Location of Salmon Farms

### 2022-03-02 marine/bivalve/bivalve_farms/STEP2_bivalve_farm_locations.Rmd
cleaned code even more and updated README

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)
library(countrycode)

@@

Based on the two datasets below, we have an existing 1010 bivalve farms. We need to add 193 points to existing 1010 (a final total of 1203 points)

@@

```{r}
maine_aq_lease <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_AQ_Leases/MaineDMR_Aquaculture_-_AQ_Leases.shp"))

@@

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - Julia Livermore: julia.livermore@dem.ri.gov
 - Current as of August 2021
 - Shellfish and kelp

@@

```{r}
rhode_island <- read_sf("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Rhode_Island/ri_aq_sites.kml")

@@

```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/North_Carolina/NC_shellfish_leases.geojson",  what = "sp") %>%
  st_as_sf()

unique(spdf$Status)

nc_active <- spdf %>%
  filter(Status %in% c("Active", "Approved")) # 423 active leases... wow!

@@

```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/CT_Aquaculture_gear/CT_aquaculture_gear.geojson",  what = "sp") %>%
  st_as_sf() %>%
  filter(Species != "kelp",
         ActivePerm == "yes")

mapview(spdf)

us_ct_bivalve_mol <- st_transform(spdf, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
us_ct_bivalve_pts <- st_centroid(us_ct_bivalve_mol)
# check
plot(st_geometry(us_ct_bivalve_mol))
plot(st_geometry(st_centroid(us_ct_bivalve_mol)), add=TRUE)

@@

```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alabama/Alabama_Shellfish_Leases.geojson",  what = "sp") %>%
  st_as_sf() %>%
  dplyr::filter(Status %in% c("Current", "Current Commercial"))

mapview(spdf)

us_al_bivalve_mol <- st_transform(spdf, crs = "+init=epsg:3035") # convert to equal area projealion to find centroids. 
us_al_bivalve_pts <- st_centroid(us_al_bivalve_mol)
# check
plot(st_geometry(us_al_bivalve_mol))
plot(st_geometry(st_centroid(us_al_bivalve_mol)), add=TRUE)

@@

## US Northeast - NY, CT, ME, NH

 Only used this for exploration
 - Regional Ocean Council 2018

Source: 
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 2020] https://www.northeastoceandata.org/data-download/

# us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))
# 
# unique(us_ne$allSpecies)
# 
# unique(fao_bivalve_spp$species)
# 
# ## need to filter allSpecies for anything with "Oyster", "Mussel", "Shellfish", "Clam", "Quahog", 
# 
# # id bivalves and active sites
# bivalve_species <- grep("Oyster|Mussel|Shellfish|Clam|Quahog", unique(us_ne$allSpecies), value=TRUE)
# 
# 
# us_ne_bivalve <- us_ne %>%
#   filter(allSpecies %in% bivalve_species) %>%
#   filter(status == "Active")
# unique(us_ne_bivalve$state)
# 
# us_ne_bivalve_mol <- st_transform(us_ne_bivalve, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
# us_ne_bivalve_pts <- st_centroid(us_ne_bivalve_mol)
# # check
# plot(st_geometry(us_ne_bivalve_mol))
# plot(st_geometry(st_centroid(us_ne_bivalve_mol)), add=TRUE)
# 
# us_ne_bivalve_pts_latlon <- st_transform(us_ne_bivalve_pts, crs = "+init=epsg:4326") 
# 
# farms_us_ne <- data.frame(do.call(rbind, st_geometry(us_ne_bivalve_pts_latlon))) # get lat lon
# names(farms_us_ne) <- c("lon", "lat") 
# farms_us_ne <- farms_us_ne %>%
#   mutate(iso3c = "USA") %>%
#   mutate(data_quality = 4) 
# 
# # all_farms_sf <- st_as_sf(farms_us_ne, coords = c("lon", "lat"), 
# #                          crs = crs(gadm_latlon), agr = "constant")
# # 
# # mapview(all_farms_sf)

>>>>

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)
library(countrycode)
# library(remotes)
# install_github("yonghah/esri2sf")
library("esri2sf")

@@

Based on the datasets below, we have an existing 1010 bivalve farms. We need to add 193 points to existing 1010 (a final total of 1203 points)

@@

```{r}
maine_aq_lease <- st_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_AQ_Leases/MaineDMR_Aquaculture_AQ_Leases.shp")

@@

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - Current as of August 2021
 - Shellfish and kelp

@@

```{r}
rhode_island <- esri2sf("https://risegis.ri.gov/hosting/rest/services/RIDEM/Marine_Fisheries_Map_Layers/MapServer/0")
mapview(rhode_island)

@@

```{r}
url <- "https://services2.arcgis.com/kCu40SDxsCGcuUWO/arcgis/rest/services/Lease_DB_View/FeatureServer/2"
nc_shellfish <- esri2sf(url)

nc_active <- nc_shellfish %>%
  filter(Status %in% c("Active", "Approved")) # 423 active leases... wow!

@@

Roadmap for polygons

Viewer: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/f8799cefbb4c4751a209710d14b9ad46/data

Search for layer you want and grab link ("CT Aquaculture Gear): https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1

Now download using esri2sf R package.

url <- "https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1"
ct_shellfish <- esri2sf(url) %>%
  st_as_sf() %>%
  filter(Species != "kelp",
         ActivePerm == "yes")

mapview(ct_shellfish)

us_ct_bivalve_mol <- st_transform(ct_shellfish, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
us_ct_bivalve_pts <- st_centroid(us_ct_bivalve_mol)
# check
plot(st_geometry(us_ct_bivalve_mol))
plot(st_geometry(st_centroid(us_ct_bivalve_mol)), add=TRUE)

@@

Roadmap to download

Viewer: https://aldcnr.maps.arcgis.com/apps/webappviewer/index.html?id=a32dad8dacd249ea86bcb80dd951a424

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/a32dad8dacd249ea86bcb80dd951a424/data

Search for layer you want and grab link ("Leases" is the layer you want): https://services7.arcgis.com/jcVUvxpWnajErgCr/arcgis/rest/services/Shellfish_Riparian_Leases_Revised/FeatureServer/0

Download using esri2sf

al_oyster <- esri2sf("https://services7.arcgis.com/jcVUvxpWnajErgCr/arcgis/rest/services/Shellfish_Riparian_Leases_Revised/FeatureServer/0") %>%
  st_as_sf() %>%
  dplyr::filter(Status %in% c("Current", "Current Commercial"))

mapview(al_oyster)

us_al_bivalve_mol <- st_transform(al_oyster, crs = "+init=epsg:3035") # convert to equal area projealion to find centroids. 
us_al_bivalve_pts <- st_centroid(us_al_bivalve_mol)
# check
plot(st_geometry(us_al_bivalve_mol))
plot(st_geometry(st_centroid(us_al_bivalve_mol)), add=TRUE)

@@

### 2022-03-02 marine/salmon/salmon_farms/STEP2_salmon_farm_locations.Rmd
cleaned code even more and updated README

# United States
Salmon aquaculture occurs along the Atlantic coast and in Puget sound.

Puget Sound data (2018 data): http://www.moldychum.com/violations-prompt-washington-state-cancel-atlantic-salmon-farm-lease/

## US Northeast
DEPRECATED:
Source:
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 17, 2020]

# us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))
# 
# 
# # id salmon and active sites
# salmon_species <- grep("Salmon|salmon", unique(us_ne$allSpecies), value=TRUE)
# 
# us_ne_salmon <- us_ne %>%
#   filter(allSpecies %in% salmon_species) %>%
#   filter(status == "Active")
# 
# us_ne_salmon_mol <- st_transform(us_ne_salmon, crs = "+init=epsg:3035") # convert to equal area projection to find centroids.
# us_ne_salmon_pts <- st_centroid(us_ne_salmon_mol)
# # check
# plot(st_geometry(us_ne_salmon_mol))
# plot(st_geometry(st_centroid(us_ne_salmon_mol)), add=TRUE)
# 
# us_ne_salmon_pts_latlon <- st_transform(us_ne_salmon_pts, crs = "+init=epsg:4326")
# 
# farms_us_ne <- data.frame(do.call(rbind, st_geometry(us_ne_salmon_pts_latlon)))
# names(farms_us_ne) <- c("lon", "lat")
# farms_us_ne <- farms_us_ne %>%
#   mutate(iso3c = "USA") %>%
#   mutate(data_quality = 2)
# 
# write_csv(farms_us_ne, here("marine/salmon/salmon_farm/data/known_farm_sites_2_4_5/farms_ne_usa.csv"))

@@

Locations of aquaculture leases in Maine, 2019. Maine Department of Marine Resources. https://dmr-maine.opendata.arcgis.com/datasets/maine::mainedmr-aquaculture-aq-leases/about

```{r}
maine_aq_lease <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_AQ_Leases/MaineDMR_Aquaculture_-_AQ_Leases.shp"))

>>>>

@@

Locations of aquaculture leases in Maine, 2019. Maine Department of Marine Resources. 
 - Viewer: https://maine.maps.arcgis.com/apps/webappviewer/index.html?id=b846cf37b1d64c988f89eafa085c8b7a
 - Download: https://dmr-maine.opendata.arcgis.com/datasets/maine::mainedmr-aquaculture-aq-leases/about
 - We will want to use regular leases only.
 - Limited purpose leases are for scientific or smaller/shorter term mariculture projects, that last up to 3 years.
 - I think this can be considered "comprehensive". 
 - Update to date as of download date
 - Has species types specified.

 - geojson API link for download: https://opendata.arcgis.com/datasets/a0b3c775cfc243a2b92df328ad85c642_2.geojson

```{r}
me_geojson <- geojson_read("https://opendata.arcgis.com/datasets/a0b3c775cfc243a2b92df328ad85c642_2.geojson",  what = "sp") %>% st_as_sf() %>%
  mutate(download_date = date, 
         data_year = year) %>%
  clean_names()
mapview(me_geojson)

#st_write(me_geojson, "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_AQ_Leases/MaineDMR_Aquaculture_AQ_Leases.shp", delete_dsn = TRUE)

maine_aq_lease <- st_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_AQ_Leases/MaineDMR_Aquaculture_AQ_Leases.shp")
